# Formula Handling

```{r start, include = FALSE}
options(width = 70)

library(parsnip)
library(workflows)
library(magrittr)
library(modeldata)
library(hardhat)
library(splines)
```

Note that, for different models, the formula given to `add_formula()` might be handled in different ways, depending on the parsnip model being used. For example, a random forest model fit using ranger would not convert any factor predictors to binary indicator variables. This is consistent with what `ranger::ranger()` would do, but is inconsistent with what `stats::model.matrix()` would do.

The documentation for parsnip models provides details about how the data given in the formula are encoded for the model if they diverge from the standard `model.matrix()` methodology. Our goal is to be consistent with how the underlying model package works.

## How is this formula used?

To demonstrate, the example below uses `lm()` to fit a model. The formula given to `add_formula()` is used to create the model matrix and that is what is passed to `lm()` with a simple formula of `body_mass_g ~ .`:

```{r pre-encoded-fit}
library(parsnip)
library(workflows)
library(magrittr)
library(modeldata)
library(hardhat)

data(penguins)

lm_mod <- linear_reg() |> 
  set_engine("lm")

lm_wflow <- workflow() |> 
  add_model(lm_mod)

pre_encoded <- lm_wflow |> 
  add_formula(body_mass_g ~ species + island + bill_depth_mm) |> 
  fit(data = penguins)

pre_encoded_parsnip_fit <- pre_encoded |> 
  extract_fit_parsnip()

pre_encoded_fit <- pre_encoded_parsnip_fit$fit

# The `lm()` formula is *not* the same as the `add_formula()` formula: 
pre_encoded_fit
```

This can affect how the results are analyzed. For example, to get sequential hypothesis tests, each individual term is tested: 

```{r pre-encoded-anova}
anova(pre_encoded_fit)
```

## Overriding the default encodings

Users can override the model-specific encodings by using a hardhat blueprint. The blueprint can specify how factors are encoded and whether intercepts are included. As an example, if you use a formula and would like the data to be passed to a model untouched:

```{r blueprint-fit}
minimal <- default_formula_blueprint(indicators = "none", intercept = FALSE)

un_encoded <- lm_wflow |> 
  add_formula(
    body_mass_g ~ species + island + bill_depth_mm, 
    blueprint = minimal
  ) |> 
  fit(data = penguins)

un_encoded_parsnip_fit <- un_encoded |> 
  extract_fit_parsnip()

un_encoded_fit <- un_encoded_parsnip_fit$fit

un_encoded_fit
```

While this looks the same, the raw columns were given to `lm()` and that function created the dummy variables. Because of this, the sequential ANOVA tests groups of parameters to get column-level p-values: 

```{r blueprint-anova}
anova(un_encoded_fit)
```

## Overriding the default model formula

Additionally, the formula passed to the underlying model can also be customized. In this case, the `formula` argument of `add_model()` can be used. To demonstrate, a spline function will be used for the bill depth:

```{r extra-formula-fit}
library(splines)

custom_formula <- workflow() |>
  add_model(
    lm_mod, 
    formula = body_mass_g ~ species + island + ns(bill_depth_mm, 3)
  ) |> 
  add_formula(
    body_mass_g ~ species + island + bill_depth_mm, 
    blueprint = minimal
  ) |> 
  fit(data = penguins)

custom_parsnip_fit <- custom_formula |> 
  extract_fit_parsnip()

custom_fit <- custom_parsnip_fit$fit

custom_fit
```

## Altering the formula

Finally, when a formula is updated or removed from a fitted workflow, the corresponding model fit is removed.

```{r remove}
custom_formula_no_fit <- update_formula(custom_formula, body_mass_g ~ species)

try(extract_fit_parsnip(custom_formula_no_fit))
```
