---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# workflows

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/tidymodels/workflows.svg?branch=master)](https://travis-ci.org/tidymodels/workflows)
[![Codecov test coverage](https://codecov.io/gh/tidymodels/workflows/branch/master/graph/badge.svg)](https://codecov.io/gh/tidymodels/workflows?branch=master)
[![R build status](https://github.com/tidymodels/workflows/workflows/R-CMD-check/badge.svg)](https://github.com/tidymodels/workflows)
<!-- badges: end -->

## What is a workflow? 

A workflow is an object that can bundle together your pre-processing, modeling, and post-processing requests. For example, if you have a `recipe` and `parsnip` model, these can be combined into a workflow. The advantages are:

 * You don't have to keep track of separate objects in your workspace.

 * The recipe prepping and model fitting can be executed using a single call to `fit()`.

 * If you have custom tuning parameter settings, these can be defined using a simpler interface when combined with [tune](https://github.com/tidymodels/tune).

 * In the future, workflows will be able to add post-processing operations, such as modifying the probability cutoff for two-class models.
 
## Installation

You can install workflows from CRAN with:

``` r
install.packages("workflows")
```


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tidymodels/workflows")
```
 
## Example

Suppose you were modeling data on cars. Say...the fuel efficiency of 32 cars. You know that the relationship between engine displacement and miles-per-gallon is nonlinear, and you would like to model that as a spline before adding it to a Bayesian linear regression model. You might have a recipe to specify the spline:

```{r spline-rec, eval = FALSE}
library(recipes)
library(parsnip)
library(workflows)

spline_cars <- recipe(mpg ~ ., data = mtcars) %>% 
  step_ns(disp, deg_free = 10)
```

and a model object:

```{r car-mod, eval = FALSE}
bayes_lm <- linear_reg() %>% 
  set_engine("stan")
```

To use these, you would generally run:

```{r car-fit, eval = FALSE}
spline_cars_prepped <- prep(spline_cars, mtcars)
bayes_lm_fit <- fit(bayes_lm, mpg ~ ., data = juice(spline_cars_prepped))
```

You can't predict on new samples using `bayes_lm_fit` without the prepped version of `spline_cars` around. You also might have other models and recipes in your workspace. This might lead to getting them mixed-up or forgetting to save the model/recipe pair that you are most interested in.

workflows makes this easier by combining these objects together:

```{r wflow, eval = FALSE}
car_wflow <- workflow() %>% 
  add_recipe(spline_cars) %>% 
  add_model(bayes_lm)
```

Now you can prepare the recipe and estimate the model via a single call to `fit()`:

```{r wflow-fit, eval = FALSE}
car_wflow_fit <- fit(car_wflow, data = mtcars)
```

You can alter existing workflows using `update_recipe()` / `update_model()` and `remove_recipe()` / `remove_model()`.
