---
title: "Getting Started"
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
output:
  knitr:::html_vignette:
    toc: yes
---
  
```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  digits = 3,
  collapse = TRUE,
  comment = "#>"
)
options(digits = 3)

library(ggplot2)

theme_set(theme_bw() + theme(legend.position = "top"))
```

This is an example of a fairly realistic interactive data analysis project to demonstrate how workflows can be used.

## Introduction

In this `bivariate` data set, there are two predictors that can be used to differentiate two classes in the outcome. 

There are three partitions of the original data: training (n = 1009), validation (n = 300), and testing (n = 710). We will work with the training set the most, use the validation set to compare models during the development process, and then use the test set once we think that we have a good algorithm for making predictions.

```{r}
library(modeldata)

# This gives us access to the 3 partitions:
# - `bivariate_train`: Training set
# - `bivariate_val`: Validation set
# - `bivariate_test`: Test set
data("bivariate")
```

Here is the training set:

```{r}
#| label: plot-data
#| message: false
#| fig-alt: A ggplot dotplot with predictor 'A' on the x axis and 'B' on the y, demonstrating
#|   a positive, linear, moderately strong correlation. Dots are colored according to
#|   two classes, red for 'One' and blue for 'Two', and the red dots tend to have slightly
#|   larger values of 'B' for given values of 'A', though the separation is far from
#|   perfect.
library(workflows)
library(ggplot2)
library(dplyr)

ggplot(bivariate_train, aes(x = A, y = B, col = Class)) + 
  geom_point(alpha = .3) + 
  coord_equal(ratio = 20)
```

Both predictors have positive values and their distributions are right-skewed. There seems to be a separation of the classes, but only when the variables are used together. For example, when the predictors are visualized individually, there is little evidence in separation of the classes.  

```{r}
#| label: plot-marginals
#| fig-alt: A pair of boxplots showing the same predictors, this time one at a time.
#|   The predictors on their own do little to separate the two outcome classes.
library(tidyr)

bivariate_train |> 
  pivot_longer(cols = c(A, B), names_to = "predictor") |> 
  ggplot(aes(x = Class, y = value)) + 
  geom_boxplot() + 
  facet_wrap(~predictor, scales = "free_y") + 
  scale_y_log10()
```

In the first plot above, the separation appears to happen linearly, and a straight, diagonal boundary might do well. We could use `glm()` directly to create a logistic regression, but we will use the `tidymodels` infrastructure and start by making a `parsnip` model object.

```{r}
#| label: glm-mod
library(parsnip)

logit_mod <-
  logistic_reg() |>
  set_engine("glm")
```

This data analysis will involve looking at a few different approaches of representing the two predictors so that we have a high-quality model. We'll walk though the thought process of this analysis as we go. This will emulate how most data analysis projects happen: an initial approach is taken and then potential steps are attempted to make improvements. There is no pre-defined blueprint to this process and the R4DS diagram summarizes the process nicely.

Since we are going to try different combinations of feature engineering and models, `workflows` are really useful since you can have one object that contains all of these operations. It helps organize your work instead of having different objects in your workspace that, at some point, have been used in pairs.

## A first set of models

The obvious place to start is by adding both predictors as-is into the model:

```{r}
#| label: simple-glm
# Create a workflow with just the model. We will add to this as we go. 
glm_workflow <-
  workflow() |>
  add_model(logit_mod)

simple_glm <-
  glm_workflow |>
  # Add both predictors in
  add_formula(Class ~ .) |>
  # Fit the model:
  fit(data = bivariate_train)
```

To evaluate this model, the ROC curve will be computed along with its corresponding AUC.

```{r}
#| label: simple-roc
#| fig-alt: A ggplot2 line plot, with 1 minus specificity on the x axis, sensitivity
#|   on the y, and a dotted line tracing the y equals x line. A solid line is shown always
#|   above the dotted line, with greatest separation from it towards the center of the
#|   plot.
library(yardstick)

simple_glm_probs <-
  predict(simple_glm, bivariate_val, type = "prob") |>
  bind_cols(bivariate_val)

simple_glm_roc <- 
  simple_glm_probs |> 
  roc_curve(Class, .pred_One)

simple_glm_probs |> roc_auc(Class, .pred_One)

autoplot(simple_glm_roc)
```

This seems reasonable. One potential issue is that the two predictors have a high degree of correlation `r round(cor(bivariate_train$A, bivariate_train$B), 3)`, and this might cause some instability in the model.

Since there are two correlated predictors with skewed distributions and strictly positive values, it might be intuitive to use their ratio instead of the pair. We'll try that next by recycling the initial workflow and just adding a different formula:

```{r}
#| label: ratios
#| fig-alt: The same plot as before, this time with a yellow line also added to the plot.
#|   The yellow line roughly traces the solid black one, except that it's y values are
#|   slightly higher than those of the solid black line towards the left portion of the
#|   plot and vice versa for the other portion.
ratio_glm <-
  glm_workflow |>
  add_formula(Class ~ I(A/B)) |> 
  fit(data = bivariate_train)

ratio_glm_probs <-
  predict(ratio_glm, bivariate_val, type = "prob") |>
  bind_cols(bivariate_val)

ratio_glm_roc <- 
  ratio_glm_probs |> 
  roc_curve(Class, .pred_One)

ratio_glm_probs |> roc_auc(Class, .pred_One)

autoplot(simple_glm_roc) + 
  geom_path(
    data = ratio_glm_roc, 
    aes(x = 1 - specificity, y = sensitivity), 
    col = "#FDE725FF"
  )
```

The original analysis shows a slight edge, but the two models are probably within the experimental noise of one another.

## More complex feature engineering

Instead of combining the two predictors, would it help the model if we were to resolve the skewness of the variables? To test this theory, one option would be to use the Box-Cox transformation on each predictor individually to see if it recommends a nonlinear transformation. The transformation can encode a variety of different functions including the log transform, square root, inverse, and fractional transformations in-between these.

This cannot be easily done via the formula interface, so a recipe is used. A recipe is a list of sequential data processing steps that are conducted before the data are used in a model. For example, to use the Box-Cox method, a simple recipe would be:

```{r}
#| label: bc-rec
#| message: false
library(recipes)

trans_recipe <- 
  recipe(Class ~ ., data = bivariate_train) |> 
  step_BoxCox(all_predictors())
```

Creating the recipe only makes an object with the instructions; it does not carry out the instructions (e.g. estimate the transformation parameter). To actually execute the recipe, we add it to our workflow with `add_recipe()` and then call `fit()`. Fitting the workflow evaluates both the model and the recipe.

```{r}
#| label: rec-trans
#| fig-alt: The same plot as above, but with another line, shown in green, added to the
#|   plot. Its y values are nearly always consistently higher than the existing black
#|   and yellow lines.
trans_glm <-
  glm_workflow |>
  add_recipe(trans_recipe) |> 
  fit(data = bivariate_train)

trans_glm_probs <-
  predict(trans_glm, bivariate_val, type = "prob") |>
  bind_cols(bivariate_val)

trans_glm_roc <- 
  trans_glm_probs |> 
  roc_curve(Class, .pred_One)

trans_glm_probs |> roc_auc(Class, .pred_One)

autoplot(simple_glm_roc) + 
  geom_path(
    data = ratio_glm_roc, 
    aes(x = 1 - specificity, y = sensitivity), 
    col = "#FDE725FF"
  ) + 
  geom_path(
    data = trans_glm_roc, 
    aes(x = 1 - specificity, y = sensitivity), 
    col = "#21908CFF"
  )
```

That is a potential, if slight, improvement.

The Box-Cox procedure recommended transformations that are pretty close to the inverse.

The model above creates a class boundary for these data:

```{r}
#| label: plot-inverse
#| fig-alt: A dotplot like that shown early on in the article, plotting the inverse of
#|   the previous two predictors and the same binary outcome variable. A boundary between
#|   the two outcome classes seems a bit more evident in this plot than before.
ggplot(bivariate_train, aes(x = 1/A, y = 1/B, col = Class)) + 
  geom_point(alpha = .3) + 
  coord_equal(ratio = 1/12)
```

The correlation between these is about the same as in the original data. It might help the model to de-correlate them, and the standard technique for this is principal component analysis. A recipe step can be added that will conduct PCA and return the score values. The scores, instead of the original predictors, can then be used in the model. PCA chases variability, so it is important to normalize the two predictors so that they have the same units. Traditionally, each column could be centered and scaled. For this reason, a step is used prior to PCA that normalizes the two predictors.

```{r}
#| label: rec-pca
pca_recipe <- 
  trans_recipe |> 
  step_normalize(A, B) |>
  step_pca(A, B, num_comp = 2)

pca_glm <-
  glm_workflow |>
  add_recipe(pca_recipe) |> 
  fit(data = bivariate_train)

pca_glm_probs <-
  predict(pca_glm, bivariate_val, type = "prob") |>
  bind_cols(bivariate_val)

pca_glm_roc <- 
  pca_glm_probs |> 
  roc_curve(Class, .pred_One)

pca_glm_probs |> roc_auc(Class, .pred_One)
```

These results are almost identical to the transformed model.

## The test set

Based on these results, the model with the logistic regression model with inverse terms is probably our best bet. Using the test set:

```{r}
#| label: test-set
#| fig-alt: An ROC curve plot like before, back to showing just one solid black line,
#|   representing what had been shown as green in the previous plot. There's quite a
#|   bit of separation between this line and the identity, showing a model that's similarly
#|   as performant as it was on the training set.
test_probs <- 
  predict(trans_glm, bivariate_test, type = "prob") |>
  bind_cols(bivariate_test)

test_roc <- 
  test_probs |> 
  roc_curve(Class, .pred_One)

# A little more optimistic than the validation set
test_probs |> roc_auc(Class, .pred_One)

autoplot(test_roc) 
```
